---
title: "Sensitivity Analysis - Sobol "
author: "Group I - Jake Eisaguirre, Yuitan Fang, Julia Parish"
date: '2022-05-19'
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=4, warning=FALSE, 
                      message=FALSE, fig.pos='!H')
```

# Sensitivity Analysis - Sobol 

This environmental model was completed as an assignment for the course, Environmental Data Science 230 | Environmental Science & Management: Modeling Environmental Systems. The goal of this assignment was to code a function to compute . This assignment focuses on developing skills to create   .

## Load Libraries
```{r, message=FALSE, warning=FALSE}

library(here)
library(tidyverse)
library(kableExtra)
library(deSolve)
library(sensitivity)

```

### Run the model for 300 years (with ODE solver) and plot the result
```{r}
# source the function

source(here("R","forestgrowthrate.R"))
dgrowthrate

# create parameter list and specify the initial size and years to run the model
initialsize = 10
years = seq(from = 1, to = 300, by = 5)
parms = list(r = 0.01, g = 2, K = 250, threshold = 50)

#apply solver
results = ode(initialsize, years, dgrowthrate, parms)
head(results)

#add meaningful names to columns of results
colnames(results) = c("year", "C")

#plot
ggplot(as.data.frame(results), aes(year,C))+geom_point()+labs(y="Forest Size (C kg)", x="Years")


```




