---
title: "Growth Model & Sensitivity Analysis - Sobol "
author: "Group I - Jake Eisaguirre, Yuitan Fang, Julia Parish"
date: '2022-05-19'
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=6, fig.height=4, fig.pos='!H')
```

# Growth Model & Sobol Sensitivity Analysis

This environmental model was completed as an assignment for the course, Environmental Data Science 230 | Environmental Science & Management: Modeling Environmental Systems. The goal of this assignment was to code a function to compute forest growth and conduct a sobol sensitivity analysis that explores how the estimated maximum and mean forest size varies. This assignment focuses on developing skills to create a model of forest growth and conduct a sobol sensitivity analysis.

### Parameters

- K = carrying capacity (C)
- r = pre-canopy closure rate
- g = post-canopy closure rate

## Load Libraries
```{r, message=FALSE, warning=FALSE}

library(here)
library(tidyverse)
library(kableExtra)
library(deSolve)
library(sensitivity)
library(purrr)
library(viridis)
```

# 1. Implement a forest growth rate model 
Forest size is measured in units of carbon (C)

```{r, model}
# source the function

source(here("R","forestgrowthrate.R"))

dgrowthrate
```

# 2. Run the model for 300 years (with ODE solver) and plot the result

## Parameters for model

- K = 250 kg C (carrying capacity)
- r = 0.01 (exponential growth rate before before canopy closure)
- g = 2 kg/year (linear growth rate after canopy closure)
- threshold = 50 kgC (canopy closure threshold)

```{r, ode}
# create parameter list and specify the initial size and years to run the model

# set parameters
K = 250
r = 0.01
g = 2
threshold = 50

initialsize <- 10
years <- seq(from = 1, to = 300, by = 1)
parms <- list(K = K, r = r, g = g,  threshold = threshold)

#apply solver
results <- ode(initialsize, years, dgrowthrate, parms)

# convert results to data frame
results <- as.data.frame(results)

#add meaningful names to columns of results
colnames(results) = c("year", "C")

```

```{r}
# view sample of df
results_sample <- head(results) 

results_table <- kable(results_sample, 
                         caption = "Sample of Forest Growth ODE Model Results") %>% 
  kable_styling(latex_options = "HOLD_position")

results_table

```


```{r, ode plot}
# plot results
model01_plot <- results %>% 
  ggplot(aes(x = year, y = C)) +
  geom_point(color = "darkgreen") + 
  labs(x = "Years", y = "Forest Size (C kg)",
       title = "Forest Growth Rate - 300 Years",
       subtitle = "K = 250 kgC, r = 0.01, g = 2 kg/yr, Canopy closure threshold = 50 kgC") +
  theme_minimal()

model01_plot
```

# 3.A. Run a sobol sensitivity analysis that explores how the estimated maximum and mean forest size (e.g maximum and mean values of C over the 300 years) varies with the pre canopy closure growth rate (r) and post-canopy closure growth rate (g) and canopy closure threshold and carrying capacity(K)

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value

```{r, parameters}

# set the number of parameters
np = 200

K = rnorm(mean = K, sd = K*0.10, n = np)
r = rnorm(mean = r, sd = r*0.10, n = np)
g = rnorm(mean = g, sd = g*0.10, n = np)
threshold = rnorm(mean = threshold, sd = threshold*0.10, n = np)

X1 = cbind.data.frame(r = r, K = K, g = g, threshold = threshold)

# repeat to calculate second set of samples
np = 200

K = rnorm(mean = K, sd = K*0.10, n = np)
r = rnorm(mean = r, sd = r*0.10, n = np)
g = rnorm(mean = g, sd = g*0.10, n = np)
threshold = rnorm(mean = threshold, sd = threshold*0.10, n = np)

X2 = cbind.data.frame(r = r, K = K, g = g, threshold = threshold)

```

```{r, sobol sensitivity}
# create sobol object and get parameters
sens_forest <- sobolSalt(model = NULL, X1, X2, nboot = 300)

sens_forestSize_df <- sens_forest$X %>% 
  as.data.frame()

sens_forestSize_df <- sens_forestSize_df %>% 
  rename(r = "V1", 
         K = "V2", 
         g = "V3", 
         threshold = "V4")
```


```{r, wrapper}
p_wrapper = function(threshold, r, g, K, initialsize, years, func) {
  
  parms <- list(threshold = threshold, r = r, g = g, K = K)
  
  forest_sensitivity <- ode(func = dgrowthrate, y = initialsize, times = years, 
                            parms = parms)
  
  forest_sensitivity <- as.data.frame(forest_sensitivity)
  
  colnames(forest_sensitivity) = c("years","C")

  # get metrics
  
  max_carbon <- max(forest_sensitivity$C)
  mean_carbon <- mean(forest_sensitivity$C)

  return(list(max_carbon=max_carbon, mean_carbon=mean_carbon))
}
```


```{r, forest senstivity results}
allresults = sens_forestSize_df %>% 
  pmap(p_wrapper, initialsize = initialsize, years = years,
       func = dgrowthrate)

allres = allresults %>% 
  map_dfr(`[`,c("max_carbon","mean_carbon"))

all_results <- pivot_longer(allres, cols = c(max_carbon, mean_carbon), names_to = "name", values_to = "carbon")
```


# 3.B. Graph the results of the sensitivity analysis as a box plot of maximum forest size and a plot of the two Sobol indices (S and T). 

```{r, box plot}
results_plot <- all_results %>% 
  ggplot(aes(y = carbon, x = name, fill = name)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  labs(x= "", y = "Carbon (kg)",
       title = "Sobol Analysis - Forest Growth",
       subtitle = "Max & mean values of Carbon over 300 years)") +
  theme_minimal() +
  theme(legend.position="none")
  
results_plot
```

```{r}
sense_mean <- sensitivity::tell(sens_forest, allres$mean_carbon)

sense_mean_S <- as.data.frame(sense_mean$S)

sense_mean_S <- sense_mean_S %>% 
  rowid_to_column(var = "parms")

sense_mean_S[1,1] <- "threshold"
sense_mean_S[2,1] <- "r"
sense_mean_S[3,1] <- "g"
sense_mean_S[4,1] <- "K"

sense_mean_T <- as.data.frame(sense_mean$T)

sense_mean_T <- sense_mean_T %>% 
  rowid_to_column(var = "parms")

sense_mean_T[1,1] <- "threshold"
sense_mean_T[2,1] <- "r"
sense_mean_T[3,1] <- "g"
sense_mean_T[4,1] <- "K"



sense_max <- sensitivity::tell(sens_forest, allres$max_carbon)

sense_max_S <- as.data.frame(sense_max$S)

sense_max_S <- sense_max_S %>% 
  rowid_to_column(var = "parms")

sense_max_S[1,1] <- "threshold"
sense_max_S[2,1] <- "r"
sense_max_S[3,1] <- "g"
sense_max_S[4,1] <- "K"


sense_max_T <- as.data.frame(sense_max$T)

sense_max_T <- sense_max_T %>% 
  rowid_to_column(var = "parms")

sense_max_T[1,1] <- "threshold"
sense_max_T[2,1] <- "r"
sense_max_T[3,1] <- "g"
sense_max_T[4,1] <- "K"

```

# 3.C. Discuss what the results of your simulation might mean for climate change impacts on forest growth
